---
title: "Plotting Diagnostics for LM and GLM with ggplot2 and ggfortify"
output: html_document
---

<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Introduction to ggfortify package}
-->

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=8, fig.height=6, fig.path='figures/lmcompat-', warning=FALSE)
```

This document explains a comparison of plotting diagnostics for LM and GLM using `ggfortify` and standard plot function. 

# Plotting Diagnostics for Linear Models 

`ggfortify` let `ggplot2` know how to interpret `lm` objects. After loading `ggfortify`, you can use `ggplot2::autoplot` function for `lm` objects. 

```{r, message = FALSE, fig.width = 5, fig.height = 6}
library(ggfortify)
m <- lm(Petal.Width ~ Petal.Length, data = iris)
autoplot(m, which = 1:6)

par(mfrow = c(3, 2))
plot(m, which = 1:6)
```

# Plotting Diagnostics for Generalized Linear Models

It also suppotgs `glm` instance.

```{r, message = FALSE, fig.width = 5, fig.height = 6}
m <- glm(Murder ~ Assault + UrbanPop + Rape,
         family = gaussian, data = USArrests)

autoplot(m, which = 1:6, ncol = 2)

par(mfrow = c(3, 2))
plot(m, which = 1:6)
```

# Specifing Plotting Options

Some properties can be changed by passing corresponding keywords. For example, `colour` keyword is for data points, `smooth.colour` is for smoothing lines and `ad.colour` is for additional auxiliary lies. Also, `ncol` and `nrow` control facet layout. Use `help(autoplot.lm)` (or `help(autoplot.*)` for any other objects) to check available options.

```{r, fig.height = 4}
autoplot(m, which = 1:6, colour = 'red',
         smooth.colour = 'black', smooth.linetype = 'dashed',
         ad.colour = 'blue',
         label.size = 3, label.n = 5, label.colour = 'blue',
         ncol = 3)
```
