---
title: "Plotting Survival Curves using ggplot2 and ggfortify"
output: html_document
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=6, fig.height=3, fig.path='figures/surv-', warning=FALSE)
```

This document explains Survival Curves related plotting using `ggplot2` and `ggfortify`. 

# Installation

First, install `ggfortify` from [github](https://github.com/sinhrks/ggfortify).

```{r, eval = FALSE, message = FALSE}
library(devtools)
install_github('sinhrks/ggfortify')
```

# Plotting with survival package

`ggfortify` let `ggplot2` know how to draw survival curves. After loading `ggfortify`, you can use `ggplot2::autoplot` function for `survfit` objects. 

```{r, message = FALSE}
library(ggplot2)
library(ggfortify)
library(survival)
d.survfit <- survfit(Surv(time, status) ~ sex, data = lung)
autoplot(d.survfit)
```

There are some options to change survival curve output. Use `help(autoplot.survfit)` (or `help(autoplot.*)` for any other objects) to check available options.

```{r, message = FALSE}
autoplot(d.survfit, surv.linetype = 'dashed', conf.int = FALSE,
         censor.shape = '*', censor.size = 5)
```

# Resources related to ggfortify

* [Plotting Time Series with ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_ts)
* [Plotting Time Series Statistics with ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_tsstats)
* [Plotting PCA/clustering results using ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_pca)
* [Plotting Survival Curves using ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_surv)
* [Plotting Probability Distributions with ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_dist)