---
title: "Plotting Multidimensional Scaling using ggplot2 and ggfortify"
output: html_document
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=6, fig.height=3, fig.path='figures/mds-', warning=FALSE)
```

This document explains MDS (Multidimensional Scaling)  plotting using `ggplot2` and `ggfortify`. 

# Installation

First, install `ggfortify` from [github](https://github.com/sinhrks/ggfortify).

```{r, eval = FALSE, message = FALSE}
library(devtools)
install_github('sinhrks/ggfortify')
```

# Before Plotting

`ggfortify` let `ggplot2` know how to interpret some popular R classes. Even though MDS functions returns `matrix` or `list` (not specific class), `ggfortify` infers background class from `list` attribute and perform `autoplot`.

**NOTE** Inference from `matrix` is not supported.

**NOTE** `ggfortify` can plot `stats::dist` instance as heatmap.

```{r}
library(ggplot2)
library(ggfortify)

data(eurodist)
d.dist <- eurodist
autoplot(d.dist)
```

# Plotting Classical (Metric) Multidimensional Scaling

`stats::cmdscale` performs Classical MDS and returns point coodinates as `matrix`, thus you can not use `autoplot` as it is. However, `stats::cmdscale` has some options to return `list`, `eig = TRUE`, `add = True` or `x.ret = True`. If any of these option is specified, `ggfortify` can infer a type and perform `autoplot`. Refer to `help(cmdscale)` to check what these options are.

```{r}
autoplot(cmdscale(d.dist, eig = TRUE))
```

Specify `label = TRUE` to plot labels.

```{r}
autoplot(cmdscale(d.dist, eig = TRUE), label = TRUE)
```

# Plotting Non-metric Multidimensional Scaling

`MASS::isoMDS` and `MASS::sammon` perform Non-metric MDS and return `list` which contains point coordinates. Thus, `autoplot` can infer its type.

**NOTE** There are some common configurable options. For details, see `help(autoplot.matrix)`.

```{r}
library(MASS)
autoplot(isoMDS(d.dist), label = TRUE, colour = 'blue',
         label.size = 5, label.colour = 'blue')
autoplot(sammon(d.dist), label = FALSE, colour = 'green')
```

# Resources related to ggfortify

* [Concepts and Basics of ggfortify](http://rpubs.com/sinhrks/basics)
* [Plotting Diagnostics for LM and GLM with ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_lm)
* [Plotting Time Series with ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_ts)
* [Plotting Time Series Statistics with ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_tsstats)
* [Plotting State Space Time Series with ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_ts_dlm)
* [Plotting PCA/clustering results using ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_pca)
* [Plotting Multidimensional Scaling using ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_mds)
* [Plotting Survival Curves using ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_surv)
* [Plotting Probability Distributions with ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_dist)
 
 　
