---
title: "Plotting Multidimensional Scaling using ggplot2 and ggfortify"
output: html_document
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=6, fig.height=3, fig.path='figures/mds-', warning=FALSE)
```

This document explains MDS (Multidimensional Scaling)  plotting using `ggplot2` and `ggfortify`. 

# Installation

First, install `ggfortify` from [github](https://github.com/sinhrks/ggfortify).

```{r, eval = FALSE, message = FALSE}
library(devtools)
install_github('sinhrks/ggfortify')
```

# Before Plotting

`ggfortify` let `ggplot2` know how to interpret some popular R classes. However, you cannot use `autoplot` directly for MDS because these function returns `matrix` or `list`, not specific class. Thus, `ggfortify` cannot define `autoplot` for MDS. You have to create an intended plot from `matrix` by yourselves. 

# Plotting Matrix

`ggfortify` defines `autoplot.matrix` to plot either heatmap or scatter plot. 

```{r, message = FALSE}
library(ggplot2)
library(ggfortify)

set.seed(1)
# heatmap
m <- matrix(rnorm(20), nc = 5)
autoplot(m)
```

To draw scatter plot, you should specify `geom = 'point'`. Target `matrix` should have 2 columns for scatter plot.

```{r}
# scatter plot
m <- matrix(rnorm(20), nc = 2)
autoplot(m, geom = 'point')
```

**NOTE** Also, `ggfortify` defines `autplot.dist` to plot `stats::dist` instance as heatmap.

```{r}
data(eurodist)
d.dist <- eurodist
autoplot(d.dist)
```

# Plotting Classical (Metric) Multidimensional Scaling

`stats::cmdscale` performs Classical MDS and returns point coodinates as `matrix`, you should call `autplot` as below. Specify `label = TRUE` to plot labels.

```{r}
autoplot(cmdscale(d.dist), geom = 'point')
autoplot(cmdscale(d.dist), geom = 'point', label = TRUE)
```

`stats::cmdscale` may return `list` depending on some options. For details, refer to `help(cmdscale)`.
In these cases, point coordinates must be explicitly passed to `autoplot`.

```{r}
autoplot(cmdscale(d.dist, eig = TRUE)$points, geom = 'point', label = TRUE)
```

# Plotting Non-metric Multidimensional Scaling

`MASS::isoMDS` and `MASS::sammon` perform Non-metric MDS. Because it returns list which contains point coordinates, it should passed explicitly to `autoplot`.

**NOTE** There is some common options can be used. For details, see `help(autoplot.matrix)`.

```{r}
library(MASS)
autoplot(isoMDS(d.dist)$points, geom = 'point', label = TRUE, colour = 'blue',
         label.size = 3, label.colour = 'blue')
autoplot(sammon(d.dist)$points, geom = 'point', label = FALSE, colour = 'green')
```

# Resources related to ggfortify

* [Plotting Diagnostics for LM and GLM with ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_lm)
* [Plotting Time Series with ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_ts)
* [Plotting Time Series Statistics with ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_tsstats)
* [Plotting PCA/clustering results using ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_pca)
* [Plotting Multidimensional Scaling using ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_mds)
* [Plotting Survival Curves using ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_surv)
* [Plotting Probability Distributions with ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_dist)
 
 　
