---
title: "Plotting Time Series with ggplot2 and ggfortify"
output: html_document
---

This document explains time series related plotting using `ggplot2` and `ggfortify`. 

# Installation

First, install `ggfortify` from [github](https://github.com/sinhrks/ggfortify).

```{r, message = FALSE}
library(devtools)
install_github('sinhrks/ggfortify')
```


# Plotting ts objects

`ggfortify` let `ggplot2` know how to interpret `ts` objects. After loading `ggfortify`, you can use `ggplot2::autoplot` function for `ts` objects. 

```{r, message = FALSE, fig.width=6, fig.height=3}
library(ggplot2)
library(ggfortify)
autoplot(AirPassengers)
```

To change line colour and line type, use `ts.colour` and `ts.linetype` options. Use `help(autoplot.ts)` (or `help(autoplot.*)` for any other objects) to check available options.

```{r, message = FALSE, fig.width=6, fig.height=3}
autoplot(AirPassengers, ts.colour = 'red', ts.linetype = 'dashed')
```

Multivariate time series will be drawn with facets. 

```{r, message = FALSE, fig.width=6, fig.height=3}
data(Canada, package = 'vars')
autoplot(Canada)
```

Specify `facet = FALSE` to draw on single axes.

```{r, message = FALSE, fig.width=6, fig.height=3}
autoplot(Canada, facet = FALSE)
```

# Plotting ts-related objects

`ggfortify` supports following time series related functions in `stats` package:

* `stl`
* `acf`, `pacf`, `ccf`
* `spec.ar`, `spec.pgram`

```{r, message = FALSE, fig.width=6, fig.height=3}
autoplot(stl(AirPassengers, s.window = 'periodic'))
```

**NOTE** With `acf` and `spec.*`, specify `plot = FALSE` to suppress default plotting outputs.

```{r, message = FALSE, fig.width=6, fig.height=3}
autoplot(acf(AirPassengers, plot = FALSE))
autoplot(spec.ar(AirPassengers, plot = FALSE))
```

# Plotting with forecast package

`ggfortify` supports `forecast` object in `forecast` package. 

```{r, message = FALSE, fig.width=6, fig.height=3}
library(forecast)
d.arima <- auto.arima(AirPassengers)
d.forecast <- forecast(d.arima, level = c(95), h = 50)
autoplot(d.forecast)
```

There are some options to change basic settings.

```{r, message = FALSE, fig.width=6, fig.height=3}
autoplot(d.forecast, ts.colour = 'green', predict.colour = 'red',
         predict.linetype = 'dashed', conf.int = FALSE)
```

# Plotting with vars package

`ggfortify` supports `varpred` object in `vars` package. 

```{r, message = FALSE, fig.width=6, fig.height=3}
library(vars)
d.vselect <- VARselect(Canada, lag.max = 5, type = 'const')$selection[1]
d.var <- VAR(Canada, p = d.vselect, type = 'const')
```

Available options are the same as `forecast`.

```{r, message = FALSE, fig.width=6, fig.height=3}
autoplot(predict(d.var, n.ahead = 50), ts.colour = 'green',
         predict.colour = 'blue', predict.linetype = 'dashed')
```

# Plotting with changepoint package

`ggfortify` supports `cpt` object in `changepoint` package. 

```{r, message = FALSE, fig.width=6, fig.height=3}
library(changepoint)
autoplot(cpt.meanvar(AirPassengers))
```

You can change some options for `cpt`.

```{r, message = FALSE, fig.width=6, fig.height=3}
autoplot(cpt.meanvar(AirPassengers), cpt.colour = 'blue', cpt.linetype = 'solid')
```