---
title: "Plotting Time Series Statistics with ggplot2 and ggfortify"
output: html_document
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=6, fig.height=3, fig.path='figures/tsstats-', warning=FALSE)
```

This document explains plotting time series statistics using `ggplot2` and `ggfortify`. 

# Installation

First, install `ggfortify` from [github](https://github.com/sinhrks/ggfortify).

```{r, eval = FALSE}
library(devtools)
install_github('sinhrks/ggfortify')
```

# Plotting time series statistics

`ggfortify` supports following time series related statistics in `stats` package:

* `stl`, `decomposed.ts`
* `acf`, `pacf`, `ccf`
* `spec.ar`, `spec.pgram`
* `cpgram` (covered by `ggcpgram`) 

```{r message = FALSE}
library(ggplot2)
library(ggfortify)
autoplot(stl(AirPassengers, s.window = 'periodic'), ts.colour = 'blue')
```

**NOTE** With `acf` and `spec.*`, specify `plot = FALSE` to suppress default plotting outputs.

```{r, fig.width = 4, fig.height = 2}
autoplot(acf(AirPassengers, plot = FALSE))
```

You can pass some options when plotting `acf` via `autoplot`. Built-in `acf` calcurates the confidence interval at plotting time and doesn't hold the result, equivalent options can be passed to `autoplot`. Following example shows to change the value of confidence interval and method (use `ma` assuming the input follows MA model).

```{r, fig.width = 4, fig.height = 2}
autoplot(acf(AirPassengers, plot = FALSE), conf.int.fill = '#0000FF', conf.int.value = 0.8, conf.int.type = 'ma')
```


```{r, fig.width = 4, fig.height = 2}
autoplot(spec.ar(AirPassengers, plot = FALSE))
```

`ggcpgram` should output the cumulative periodogram as the same as `cpgram`. Because `cpgram` doesn't have return value, we cannot use `autoplot(cpgram(...))`.

```{r, message = FALSE, fig.width=3, fig.height=3}
ggcpgram(arima.sim(list(ar = c(0.7, -0.5)), n = 50))
```

`ggtsdiag` should output the similar diagram as `tsdiag`.

```{r, message = FALSE, fig.width = 5, fig.height = 5}
library(forecast)
ggtsdiag(auto.arima(AirPassengers))
```

`gglagplot` is for `lag.plot`.

```{r, message = FALSE, fig.width = 4, fig.height = 4}
gglagplot(AirPassengers, lags = 4)
```

# Resources related to ggfortify

* [Plotting Time Series with ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_ts)
* [Plotting Time Series Statistics with ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_tsstats)
* [Plotting PCA/clustering results using ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_pca)
* [Plotting Survival Curves using ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_surv)
* [Plotting Probability Distributions with ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_dist)