---
title: "Plotting State Space Time Series with ggplot2 and ggfortify"
output: html_document
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=6, fig.height=3, fig.path='figures/tsss-', warning=FALSE)
```

This document explains state space time series related plotting using `ggplot2` and `ggfortify`. 

# Installation

First, install `ggfortify` from [github](https://github.com/sinhrks/ggfortify).

```{r, eval = FALSE}
library(devtools)
install_github('sinhrks/ggfortify')
```

# Plotting with `{dlm}` package

`ggfortify` let `ggplot2` know how to interpret `{dlm}` instances. After loading `ggfortify`, you can use `ggplot2::autoplot` to plot them. 

```{r, message = FALSE}
library(ggfortify)
```

`autoplot` draws both original and fitted time series, because `dlm::dlmFilter` contains them. 

```{r, message=FALSE, warning=FALSE}
library(dlm)
form <- function(theta){
  dlmModPoly(order = 1, dV = exp(theta[1]), dW = exp(theta[2]))
}

model <- form(dlmMLE(Nile, parm = c(1, 1), form)$par)
filtered <- dlmFilter(Nile, model)

autoplot(filtered)
```

You can specify some options to change how plot looks, such as `ts.linetype` and `fitted.colour`. Use `help(autoplot.tsmodel)` (or `help(autoplot.*)` for any other objects) to check available options.

```{r}
autoplot(filtered, ts.linetype = 'dashed', fitted.colour = 'blue')
```

`ggfortify` can plot `dlm::dlmSmooth` instance which returns a `list` instance using using class inference. Note that only smoothed result is plotted because `dlm::dlmSmooth` doesn't contain original `ts`.

```{r}
smoothed <- dlmSmooth(filtered)
class(smoothed)
autoplot(smoothed)
```

To plot original `ts`, filtered result and smoothed result, you can use`autoplot` as below. When `autoplot` accepts `ggplot` instance via `p` option, continuous plot is drawn on the passed `ggplot`. 

```{r}
p <- autoplot(filtered)
autoplot(smoothed, ts.colour = 'blue', p = p)
```

# Plotting with `{KFAS}` package

You can use `autoplot` in almost the same manner as `{dlm}`. Note that `autoplot` draws smoothed result if it exists in `KFAS::KFS` instance, and `KFAS::KFS` contains smoothed result by default. 

```{r, message=FALSE, warning=FALSE}
library(KFAS)
model <- SSModel(
  Nile ~ SSMtrend(degree=1, Q=matrix(NA)), H=matrix(NA)
)
 
fit <- fitSSM(model=model, inits=c(log(var(Nile)),log(var(Nile))), method="BFGS")
smoothed <- KFS(fit$model)
autoplot(smoothed)
```

If you want filtered result, specify `smoothing='none'` when calling `KFS`. For details, see `help(KFS)`.

```{r}
filtered <- KFS(fit$model, filtering="mean", smoothing='none')
autoplot(filtered)
```

Also, `KFAS::signal` will retrieve specific state from `KFAS::KFS` instance. The result will be a `list` which contains the retrieved state as `ts` in `signal` attribute. `ggfortify` can autoplot it using class inference. 

```{r}
trend <- signal(smoothed, states="trend")
class(trend)
```

Because `signal` is a `ts` instance, you can use `autoplot` and `p` option as the same as `dlm::dlmSmooth` example.

```{r}
p <- autoplot(filtered)
autoplot(trend, ts.colour = 'blue', p = p)
```

# Resources related to ggfortify

* [Concepts and Basics of ggfortify](http://rpubs.com/sinhrks/basics)
* [Plotting Diagnostics for LM and GLM with ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_lm)
* [Plotting Time Series with ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_ts)
* [Plotting Time Series Statistics with ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_tsstats)
* [Plotting State Space Time Series with ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_ts_dlm)
* [Plotting PCA/clustering results using ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_pca)
* [Plotting Multidimensional Scaling using ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_mds)
* [Plotting Survival Curves using ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_surv)
* [Plotting Probability Distributions with ggplot2 and ggfortify](http://rpubs.com/sinhrks/plot_dist)
 
 　
